---
title: "Project 2"
author: "Ethan Lewis"
date: "10/31/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(ggplot2)
```

```{r}
source("http://www.openintro.org/stat/data/cdc.R")
# LOAD IN DATA
```

```{r}
 get_logitplot <- function(x, y, xname, bins, formulahere){

    # Step 1 : Create storage space needed for the loop
    nbins <- length(bins)
    probs.each <-NULL

    # Step 2: Obtain the probability
    for(i in 1:nbins){
      if( i < nbins){
        scores.in <- which(x< bins[i+1] & x >= bins[i])
      } else{
        scores.in <- which(x> bins[i])
      }
      numerator  <- length(which(y[scores.in]==1))
      denominator      <- length(which(y[scores.in]==0))
      probs.each <- c(probs.each,ifelse(numerator>0 & denominator>0,numerator/denominator,0))
    }

    #  Step 3: Convert to the log Odds
    log.RR.each <- log(probs.each)

    to.remove <- which(log.RR.each=="-Inf")
    log.RR.each <-log.RR.each[-to.remove]
    bins <- bins[-to.remove]

    dataHere <-data.frame(c(bins), c(log.RR.each))

    ggplot(dataHere, aes(x =bins, y = log.RR.each)) + geom_point() + geom_smooth(method = "lm", formula = formulahere, se = FALSE)+labs(x = xname, y = "Log Odds")
 }
# CREATE get_logitplot FUNCTION
```

## Executive Summary

No Code

## Section 1: Introduction

No Code

## Section 2: Exploratory Data Analysis

```{r, results='hide'}
summary(cdc)
# SUMMARY OF cdc 
```

```{r}
knitr::kable(table(cdc$exerany), col.names=c("Exercise", "Count"), caption = "'exerany' Distribution")
# exerany COUNT TABLE
```

```{r}
heightEDA <- get_logitplot(x=cdc$height , y=cdc$exerany, xname = "Height (in.)", bins = seq(from = 48, to = 93, by = 3), formula = y ~ x) + labs(title="Figure 2.1")
# EDA height
```

```{r}
weightEDA <- get_logitplot(x=cdc$weight , y=cdc$exerany, xname = "Weight (lbs.)", bins = seq(from = 68, to = 500, by = 10), formula = y ~ poly(x,2)) + labs(title="Figure 2.2")
# EDA weight
```

```{r}
wtdesireEDA <- get_logitplot(x=cdc$wtdesire , y=cdc$exerany, xname = "Desired Weight (lbs.)", bins = seq(from = 68, to = 680, by = 10), formula = y ~ poly(x,2)) + labs(title="Figure 2.3")
# EDA wtdesire
```

```{r}
ageEDA <- get_logitplot(x=cdc$age , y=cdc$exerany, xname = "Age", bins = seq(from = 18, to = 99, by = 1), formula = y ~ x) + labs(title="Figure 2.4")
# EDA age
```

```{r}
gridExtra::grid.arrange(heightEDA,weightEDA,wtdesireEDA,ageEDA, ncol = 2)
```

```{r, fig.asp=0.6}
par(mfrow = c(1,2))

mosaicplot(gender~exerany, data = cdc, xlab = "Gender", ylab = "Exercise in Past Month" , main = "Figure 2.5")

mosaicplot(smoke100~exerany, data = cdc, xlab = "Previously Smoked 100 Cigarettes", ylab = "Exercise in Past Month" , main = "Figure 2.6")
# EDA gender AND smoke100
```

```{r, fig.asp=0.5}
ggplot(cdc, aes(x= exerany, fill = factor(hlthplan))) + geom_histogram(bins=10, color = "black") +  facet_wrap( ~ hlthplan, ncol=2) + labs(title="Figure 2.7: Health Coverage", x="Exercise in Past Month") + scale_x_continuous(breaks=c(0,1))
# EDA hlthplan
```

```{r}
levels(cdc$genhlth)[levels(cdc$genhlth)=="very good"] <- "excellent"
# COMBINE 'excellent' AND 'very good' LEVELS AS 'excellent'

mosaicplot(genhlth~exerany, data = cdc, xlab = "General Health Self Evaluation", ylab = "Exercise in Past Month" , main = "Figure 2.8")
# NEW EDA genhlth
```

## Section 3: Modeling 

$Y_{i}$ ~ Bernoulli($\pi_{i}$)

$$log\left(\frac{\pi_i}{1-\pi_i}\right) = \beta_0 + \beta_1(Height_i) + \beta_2(Weight_i) + \beta_3 (Weight_i^{2}) + \beta_4(DesiredWeight_i) + \beta_5(DesiredWeight_i^{2}) + \beta_6(Age_i) +$$
$$\beta_7(GenderF_i) + \beta_8(HealthCoverage_i) + \beta_9(Smoke100_i) + \beta_{10}(GoodHealth_i) + \beta_{11}(FairHealth_i) + \beta_{12}(PoorHealth_i)$$

```{r, results='hide'}
model1 <- glm(exerany~height+poly(weight,2)+poly(wtdesire,2)+age+gender+hlthplan+smoke100+genhlth, data = cdc, family = "binomial")
summary(model1)
# FIT FULL MODEL, model1, AND RUN SUMMARY
```

```{r, results='hide'}
BIC(model1)
# model1 BIC
```

$Y_{i}$ ~ Bernoulli($\pi_{i}$)

$$log\left(\frac{\pi_i}{1-\pi_i}\right) = \beta_0 + \beta_1(Height_i) + \beta_2(Weight_i) + \beta_3(Weight_i^2) + \beta_4(Age_i) +$$
$$\beta_5(HealthCoverage_i) + \beta_6(GoodHealth_i) + \beta_7(FairHealth_i) + \beta_8(PoorHealth_i)$$

```{r, results='hide'}
model2 <- glm(exerany~height+weight+I(weight^2)+age+hlthplan+genhlth, data = cdc, family = "binomial")
summary(model2)
# FIT FULL MODEL, model1, AND RUN SUMMARY
```

```{r, results='hide'}
BIC(model2)
# model2 BIC
```

(1) $H_O:$ Model 1, the full model, is not a better fit 

    $H_A:$ Model 1, the full model, is a better fit 

(2) $\widehat{D_{Model1}}$ = 21315 where $p_{Model1}$ = 13

    $\widehat{D_{Model2}}$ = 21322 where $p_{Model2}$ = 9

(3) G = 21322 - 21315 = 7

(4) G ~ $\chi^2(4)$

```{r, results='hide'}
pchisq(7, df = 4, lower.tail = FALSE)
# CALCULATE P-VALUE
```

$$log\left(\frac{\hat{\pi_i}}{1-\hat{\pi_i}}\right) = -3.146 + 0.07545(Height_i) - 0.001794(Weight_i) - 0.000007715(Weight_i^2) - 0.006703(Age_i) +$$ 
$$0.4392(HealthCoverage_i) - 0.5714(GoodHealth_i) - 0.9747(FairHealth_i) - 1.534(PoorHealth_i)$$


```{r, results='hide', message=FALSE}
confint(model2, level = 0.95)
# 95% CONFIDENCE INTERVALS FOR MODEL 2
```

- $\beta_0$: (-3.80561, -2.48915)

- $\beta_1(Height_i)$: (0.06542, 0.085501)

- $\beta_2(Weight_i)$: (-0.006364, 0.002833)

- $\beta_3(Weight_i^2)$: (-0.00001932, 0.000003697)

- $\beta_4(Age_i)$: (-0.008716, -0.0046884)

- $\beta_5(HealthCoverage_i)$: (0.34246, 0.535313)

- $\beta_6(GoodHealth_i)$: (-0.64672, -0.49592)

- $\beta_7(FairHealth_i)$: (-1.0791, -0.87012)

- $\beta_8(PoorHealth_i)$:(-1.69902, -1.36959)

## Section 4: Prediction

$$\hat{\pi_1} = \frac{e^{-3.146 + 0.07545(72) - 0.001794(190) - - 0.000007715(36100)- 0.006703(26) + 0.4392 - 0.5714}}{1+ e^{-3.146 + 0.07545(72) - 0.001794(190) - - 0.000007715(36100)- 0.006703(26) + 0.4392 - 0.5714}}$$

$$\hat{\pi_1} = 0.87186$$

```{r}
set.seed(279)
preds <- data.frame(predict(model2))
preds[preds > 0] <- 1
preds[preds < 0] <- 0
# CREATE preds 
```

```{r, results='hide'}
table("Prediction" = preds$predict.model2., "Actual" = cdc$exerany)
# preds CONFUSION MATRIX
```

$CER_{total} = \frac{4553+425}{20000} = 24.89\%$

$CER_{NoExercise} = \frac{4553}{5086} = 89.52\%$

$CER_{Exercise} = \frac{425}{14914} = 2.85\%$

## Section 5: Conclusion

No Code 




